---
title: "babette demo"
author: "Richel J.C. Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{babette demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates how to use `babette`.

First, load the library:

```{r load_babette, results='hide', warning=FALSE, error=FALSE, message=FALSE}
library(babette)
```

The following examples show:

 * one alignment
 * two alignments
 * two alignments with a fixed crown age

Of each example, it is shown:

 * How to let `babette` run BEAST2
 * How the plot the posterior estimates
 * How to show the effective sample sizes (ESS)
 
In all cases, this is done for a short MCMC chain length of 10K:

```{r}
sample_interval <- 1000
mcmc = create_mcmc(
  chain_length = 10000, 
  store_every = sample_interval
)
```

For a reliable inference, use an ESS of at least 200.

## One alignment

```{r cache=TRUE}
out <- bbt_run(
  get_babette_path("anthus_aco.fas"),
  mcmc = mcmc
)
```

```{r}
library(ggplot2)
p <- ggplot(
  data = out$estimates,
  aes(x = Sample)
) 
p + geom_line(aes(y = TreeHeight), color = "green")
p + geom_line(aes(y = YuleModel), color = "red")
p + geom_line(aes(y = birthRate), color = "blue")

```

Effective sample sizes, with 20% burn-in removed:

```{r}
traces <- remove_burn_ins(
  traces = out$estimates, 
  burn_in_fraction = 0.2
)
esses <- t(calc_esses(traces, sample_interval = sample_interval))
colnames(esses) <- "ESS"
knitr::kable(esses)
```

Summary statistics:
  
```{r}
sum_stats <- t(calc_summary_stats(traces$posterior, sample_interval = sample_interval))
colnames(sum_stats) <- "Statistic"
knitr::kable(sum_stats)
```

Phylogenies:

```{r fig.width=7, fig.height=7}
plot_densitree(out$anthus_aco_trees, width = 2)
```

## Two alignments

```{r cache=TRUE}
out <- bbt_run(
  get_babette_paths(c("anthus_aco.fas", "anthus_nd2.fas")),
  mcmc = mcmc
)
```

```{r}
p <- ggplot(
  data = out$estimates,
  aes(x = Sample)
) 
p + geom_line(aes(y = TreeHeight.aco), color = "green") +
   geom_line(aes(y = TreeHeight.nd2), color = "lightgreen")
p + geom_line(aes(y = YuleModel.aco), color = "red") +
  geom_line(aes(y = YuleModel.nd2), color = "pink")
p + geom_line(aes(y = birthRate.aco), color = "blue") + 
  geom_line(aes(y = birthRate.nd2), color = "cyan")

```

Effective sample sizes, with 20% burn-in removed:

```{r}
traces <- remove_burn_ins(traces = out$estimates, burn_in_fraction = 0.2)
esses <- t(calc_esses(traces, sample_interval = sample_interval))
colnames(esses) <- "ESS"
knitr::kable(esses)
```

Phylogenies:

```{r fig.width=7, fig.height=7}
plot_densitree(out$anthus_aco_trees, width = 2)
```

```{r fig.width=7, fig.height=7}
plot_densitree(out$anthus_nd2_trees, width = 2)
```

## Two alignments, fixed crown age

```{r cache=TRUE}
out <- bbt_run(
  get_babette_paths(c("anthus_aco.fas", "anthus_nd2.fas")),
  mcmc = mcmc,
  posterior_crown_age = 15
)
```

```{r}
p <- ggplot(
  data = out$estimates,
  aes(x = Sample)
) 
p + geom_line(aes(y = TreeHeight.aco), color = "green") +
   geom_line(aes(y = TreeHeight.nd2), color = "lightgreen")
p + geom_line(aes(y = YuleModel.aco), color = "red") +
  geom_line(aes(y = YuleModel.nd2), color = "pink")
p + geom_line(aes(y = birthRate.aco), color = "blue") + 
  geom_line(aes(y = birthRate.nd2), color = "cyan")

```

Effective sample sizes, with 20% burn-in removed:

```{r}
traces <- remove_burn_ins(traces = out$estimates, burn_in_fraction = 0.2)
esses <- t(calc_esses(traces, sample_interval = sample_interval))
colnames(esses) <- "ESS"
knitr::kable(esses)
```

Phylogenies:

```{r fig.width=7, fig.height=7}
plot_densitree(out$anthus_aco_trees, width = 2)
```

```{r fig.width=7, fig.height=7}
plot_densitree(out$anthus_nd2_trees, width = 2)
```
