---
title: "Nested Sampling"
author: "Richel J.C. Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Nested Sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to use the Nested Sampling approach, as
described in [1]

## Setup

```{r}
library(babette)
```

Check if BEAST2 is installed. If not, install it

```{r}
if (!file.exists(get_default_beast2_path()))
{
  print("Installing BEAST2 to default location")
  install_beast2()
}
```

```{r}
if (!mauricer::mrc_is_installed("NS"))
{
  print("Installing BEAST2 package 'NS'")
  mauricer::mrc_install("NS")
}
```

## Do the run

```{r}
beast2_input_file <- tempfile(fileext = ".xml")
beautier::create_beast2_input_file(
  input_filenames = get_babette_path("anthus_aco_sub.fas"),
  output_filename = beast2_input_file,
  mcmc = create_mcmc_nested_sampling(chain_length = 10000)
)
```

```{r}
out <- bbt_run(
  fasta_filenames = get_babette_path("anthus_aco_sub.fas"),
  mcmc = create_mcmc_nested_sampling(chain_length = 20000, store_every = 1000),
  beast2_input_filename = beast2_input_file,
  beast2_path = beastier::get_default_beast2_bin_path()
)
```

Show the results, where `NSlikelihood` is the Nested-Sampling likelihood:

```{r}
knitr::kable(out$estimates)
```

## References

 * [1] Maturana, P., Brewer, B. J., Klaere, S., & Bouckaert, R. (2017).
   Model selection and parameter inference in phylogenetics
   using Nested Sampling. arXiv preprint arXiv:1703.05471.
