---
title: "babette: Step by Step"
author: "Richel J.C. Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{babette: Step by Step}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This demo shows how to run the `babette` pipeline step by step.

The `babette` demo shows how to use `babette` at its highest level:
all intermediate steps are ignored, and all intermediate files are put
in a temporary folder.

```{r load_babette, results='hide', warning=FALSE, error=FALSE, message=FALSE}
library(babette)
```

In all cases, this is done for a short MCMC chain length of 10K:

```{r}
sample_interval <- 1000
mcmc <- create_mcmc(
  chain_length = 10000, 
  store_every = sample_interval
)
```

## Create a BEAST2 input file

This step is commonly done using BEAUti. 
With `babette`, this can be done as follows:

```{r}
beast2_input_file <- tempfile(fileext = ".xml")
create_beast2_input_file(
  get_babette_path("anthus_aco.fas"),
  output_filename = beast2_input_file,
  mcmc = mcmc
)
testit::assert(file.exists(beast2_input_file))
```

## Display (part of) the BEAST2 input file

```{r}
print(head(readLines(beast2_input_file)))
```

This file can both be loaded by BEAUti and be used by BEAST2.

## Run MCMC

This step is commonly done using BEAST2 from the command-line or using its GUI. 
With `babette`, this can be done as follows:

```{r}
log_filename <- tempfile(fileext = ".log")
trees_filename <- tempfile(fileext = "trees")
state_filename <- tempfile(fileext = ".xml.state")
run_beast2(
  input_filename = beast2_input_file,
  output_log_filename = log_filename,
  output_trees_filenames = trees_filename,
  output_state_filename = state_filename
)
testit::assert(file.exists(log_filename))
testit::assert(file.exists(trees_filename))
testit::assert(file.exists(state_filename))

```

## Display (part of) the BEAST2 output files

The `.log` file contains the model parameters and parameter estimates:

```{r}
print(head(readLines(log_filename)))
print(tail(readLines(log_filename)))
```

The `.trees` file contains the alignment, taxa and posterior trees:

```{r}
print(head(readLines(trees_filename)))
print(tail(readLines(trees_filename)))
```

The `.xml.state` file contains the final state of the MCMC run and the
MCMC operator acceptances thus far: 

```{r}
print(head(readLines(state_filename)))
print(tail(readLines(state_filename)))
```

## Parse output

This step is commonly done using Tracer. 
With `babette`, this can be done as follows.

Parsing `.log` file to obtain the parameter estimates:

```{r}
knitr::kable(head(parse_beast_log(log_filename)))
```

Parsing `.trees` file to obtain the posterior phylogenies:

```{r fig.width = 7, fig.height = 7}
plot_densitree(parse_beast_trees(trees_filename))
```

Parsing `.xml.state` file to obtain the MCMC operator acceptances:

```{r}
knitr::kable(head(parse_beast_state_operators(state_filename)))
```
